\documentclass[12pt]{article}

\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{lmodern}
\usepackage{hyperref}

\title{Ardana Rollups System Architecture}
\author{Ardana Labs}


\begin{document}


\maketitle


Ardana Rollups are a general purpose ZK rollup layer 2 scalability solution for Cardano. ZK rollups are a mechanism by which transactions are processed off-chain and the results of those transactions are published onto the blockchain using zero knowledge proofs. More precisely, ZK rollups use zkSNARKS, which are in other words zero knowledge succinct non-interactive arguments of knowledge. By moving data and processing off-chain, ZK rollups allow for higher throughput. By using a secure, decentralized layer 1 blockchain and zkSNARKS, ZK rollups inherit certain security properties of the layer 1 solution, namely that we need not assume that any party is trustworthy, and nobody is able to violate the protocol and cause incorrect results to be produced on-chain. Additionally, Ardana Rollups are designed to have the security property that there is no single point of failure, meaning that there is no individual or organization whose ongoing participation is required for the proper, intended functioning of the solution.

ZK rollups are an increasingly popular choice of layer 2 solution on Ethereum, with multiple choices of ZK rollup solution available for use, including zkSync, StarkEx, and Loopring. \cite{ethworks-20} In contract, we know of no prior art for a ZK rollup based layer 2 solution on Cardano. However, we believe that ZK rollups will be an important part of the Cardano ecosystem moving forward, just as they are in Ethereum today.

This is the system architecture document for Ardana Rollups. This defines the key components of Ardana Rollups and their interactions with each other and the outside world, as well as the rationale for those design decisions. This is different from a whitepaper or a technical spec in that it does not attempt to provide enough information to implement Ardana Rollups. Instead, you can think of it as a sketch. Technical specs will fill in the picture in much greater detail, sufficient to guide implementation, and a technical whitepaper will describe Ardana Rollups in a form more suitable for general consumption.


\begin{thebibliography}{1}


	\bibitem{ethworks-20}
		Ethworks. \textit{Zero-Knowledge Blockchain Scalability}. Ethworks Reports, 2020. \url{https://ethworks.io/assets/download/zero-knowledge-blockchain-scaling-ethworks.pdf}


\end{thebibliography}



\end{document}
